Phenotypic evolution is more constrained in simpler food webs

### 2019-01-09 manuscript.Rmd
update fig. 3 with arrows and contours

\noindent 1. University of British Columbia, Department of Zoology, Vancouver, BC V6T 1Z4, Canada;

@@

*Keywords*: fitness landscape, host-parasitoid, natural selection, G-matrix, community context, ecological networks.

@@

@@

The adaptive landscape provides a powerful framework for understanding how natural selection has shaped the evolution of biodiversity ---from genes to phenotypes to species [@Wright1931; @Simpson1944; @Arnold2001]. More than a metaphor, the adaptive landscape links quantitative genetic and phenotypic variation to evolution by natural selection [@Lande1979; @Arnold1984applications; @Arnold1984theory]. Ecological interactions often play a key role in shaping adaptive landscapes, as evidenced by the role of antagonistic [@Schluter2000; @Abrams2000] and mutualistic [@Bronstein2006] interactions in driving evolutionary change.<!--, often play a key role in shaping the adaptive landscape and phenotypic evolution in multiple traits [; ;]. For example, competition for shared prey in threespine stickleback has driven their adaptive divergence into ecotypes with specialized body and gill morphologies [@Schluter2000, other papers]. Similarly, the elongated, red corollas of many flowering species have evolved in response to the efficient pollination of hummingbirds [cite].--> Although there is clear evidence that pairwise interactions can shape the adaptive landscape, we do not know how the adaptive landscape is shaped by the community context [@McPeek2017; @terHorst2018]. Resolution on how the community context shapes phenotypic evolution is urgently needed though, given the rapid impacts of climate change on ecological communities [@Scheffers2016].

Ecological networks, such as food webs describing who-eats-whom in ecological communities, provide an explicit representation of the community context [@Bascompte2014; @McCann2012].<!--These networks describe the interdependency of populations within an ecological community, providing an effective framework for predicting how the strength and organization of ecological interactions (network structure) affect population and community dynamics [@Bascompte2014; @McCann2012]. Nevertheless, these networks often lack quantitative information on how ecological interactions influence selection on phenotypic traits. Thus, it remains unclear how network structure shapes adaptive landscapes of interacting populations. This represents a major barrier in our current ability to predict how changes in community context will affect phenotypic evolution.--> Here, we integrate ecological networks and adaptive landscapes to understand how community context constrains evolutionary change [@Arnold1992]. Different aspects of evolutionary constraints can be inferred by quantifying the slope and curvature of the adaptive landscape. For example, the slope is determined by directional selection gradients acting on each phenotypic trait and influences the trajectory of evolutionary change [@Lande1979;@Arnold1992]. Evolutionary trajectories become more constrained with an increase in the number of traits under selection, as this diminishes the number of optimal solutions [@Arnold2003]. The curvature of the adaptive landscape can also constrain evolution through its indirect effect on genetic constraints [@Arnold1992; @Hansen2008]. Genetic constraints are largely governed by a population's **G**-matrix ---the additive genetic variances and covariances between traits [@Hansen2008]. In general, genetic constraints will increase with the number of traits under directional or stabilizing selection, as this will decrease the additive genetic variance in those traits [@Hansen2008]. Genetic constraints may also increase with the number of trait combinations under correlational selection, as this type of selection decreases the evolutionary independence of traits [@Hansen2008]. If we want to predict how community context constrains evolutionary change, we must understand how ecological networks shape the adaptive landscape.

Global change is simplifying the structure of ecological communities, which may influence evolutionary constraints in a number of ways. For example, in a multitrophic community or food web, if different consumers impose directional selection on different traits of a shared resource, then more complex food webs may constrain evolution by increasing the number of traits under selection. Alternatively, if consumers impose selection on different values of a trait, then their selective effects would cancel each other out in more complex food webs. To examine these different possibilities (among others), <!--To examine the effect of network structure on adaptive landscapes, -->we conducted a field experiment that manipulated the complexity of trophic interactions (hereafter food-web complexity) associated with an abundant insect herbivore (*Iteomyia salicisverruca*; Family Cecidomyiidae)(fig. \ref{fig:Conceptual}). The larvae of this herbivore induce tooth-shaped galls when they feed on the developing leaves of willow trees [*Salix* sp., @Russo2006]. These galls protect larva from attack by generalist predators (e.g. ants, spiders), but they suffer high mortaltiy from egg and larval parasitoids [@Barbour2016]. We manipulated food-web complexity by either excluding the guild of larval parasitoids (simple food web) or allowing both egg and larval parasitoids to impose selection on gall midge traits (complex food web; fig. \ref{fig:Conceptual}). <!--Food-web complexity may influence the adaptive landscape in a number of ways. For example, if each parasitoid imposes directional selection on a different trait, then more complex food webs may constrain evolution by increasing the number of traits under selection. Alternatively, if parasitoids impose selection on different values of a trait, then their selective effects would cancel each other out in more complex food webs. To examine these different possibilities (among others), -->We then applied modern statistical methods to quantify the effect of food-web complexity on the slope and curvature of the gall midge's adaptive landscape. Taken together, our study gives insight to how local extinctions, and concomitant decreases in food-web complexity, may constrain the evolution of interacting populations.

@@

We collected data on three different traits that we expected to influence larval survival based on our previous work [@Barbour2016] and others work with gall midges [@Weis1983; @Heath2018]. First, we measured gall diameter as the size of each gall chamber to the nearest 0.01 mm at its maximum diameter (perpendicular to the direction of plant tissue growth). Our previous work has shown that larger galls are associated with higher survival [@Barbour2016]. Second, we measured clutch size by counting the number of chambers in each gall [@Weis1983; @Heath2018]. All larvae collected from the same multi-chambered gall were scored with the same clutch size. Third, we measured oviposition preference as the density of larvae observed on a plant in an independent survey. We did this randomly sampling five branches per tree and counting the number of individual gall chambers (number of larvae). We then converted these counts to a measure of larval density per 100 shoots by counting the number of shoots on the last branch we sampled. All larvae collected from the same plant were scored with the same oviposition preference. Measuring larval densities on plants in the field is a common method for measuring oviposition preference [@Gripenberg2010]; however, caution must be taken in inferring 'preference' as larval densities can be influenced by processes other than preference [@Singer1986]. Fortunately, a couple of features of our study system suggest that larval densities may be a good proxy for oviposition preference. For example, since our data comes from a randomized placement of plant genotypes in a common garden, there is no consistent bias in which plant genotypes females are exposed to while searching for oviposition sites. Also, egg predation is a minor source of mortality for galling insects in general [@Hawkins1997]; therefore, we do not expect any prior egg predation to bias our estimates of oSimilar patterns in larval parasitoids has been observed for other a related gall midge species [@Weis1982]Similar patterns in larval parasitoids has been observed for other a related gall midge species [@Weis1982]bserved larval densities.

@@

The indirect effects of selection on genetic constraints is governed by the curvature of the adaptive landscape, which in our study corresponds to:
<!--The curvature (**C**) of the adaptive landscape is a matrix that is determined by directional, quadratic, and correlational selection gradients, such that \boldmath$\text{C}=\gamma - \beta \beta^\text{T}$, or more explicitly:-->

@@

All analyses and visualizations were conducted in R [@R2018]. Unless otherwise noted, we report mean estimates of trait-fitness relationships and selection gradients with 95% confidence intervals in brackets. Note that for visualizing the adaptive landscape weSimilar patterns in larval parasitoids has been observed for other a related gall midge species [@Weis1982] restrict trait axes to $\pm 1$ SD of the mean trait value. This emphasizes the fact that we can only reliably estimate the shape of the adaptive landscape near the mean phenotype of the population [@Arnold2001]. We also plot mean larval survival on a natural log scale to accurately reflect the shape of the adaptive landscape [@Arnold2003].

@@

## Food-web complexity reduces selective constraints

\indent We found that more gall midge traits experienced directional selection in the simple (3 of 3) vs. complex food webs (1 of 3)(table \ref{Table:Gradients}). For example, we observed directional selection for smaller clutch sizes in the simple food web, but there was no evidence of selection acting on this trait in the complex food web (table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}C). This absence of selection appeared to be a result of conflicting selection pressures imposed by each guild of parasitoids (fig. \ref{fig:UV_Landscape}C). Specifically, when we subset our data to focus on differences between parasitoid guilds, we found that larval parasitoids actually impose directional selection for larger clutch sizes ($\beta_{\text{Clutch}}$=
`r round(filter(tidy_egglarval_grads, type=="Complex", term=="sc.log.Clutch")$mean,2)` 
[`r round(filter(tidy_egglarval_grads, type=="Complex", term=="sc.log.Clutch")$conf.low,2)`,
`r round(filter(tidy_egglarval_grads, type=="Complex", term=="sc.log.Clutch")$conf.high,2)`]). In the simple food web, we also observed clear evidence of directional selection for midges to avoid ovipositing on plants with high densities of conspecifics (table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}B); however, this relationship was weaker in complex food webs (table \ref{Table:Gradients}). This was likely a result of larval parasitoids imposing greater mortality on egg parasitoids at high gall midge densities (see Selection on the extended phenotype of egg parasitoids section), and thus a less than additive effect on gall midges. Chamber diameter experienced positive directional selection in both food-web treatments, but selection was `r round(((filter(tidy_foodweb_grads, type=="Complex", term=="sc.Diam")$mean - filter(tidy_foodweb_grads, type=="Simple", term=="sc.Diam")$mean) / filter(tidy_foodweb_grads, type=="Simple", term=="sc.Diam")$mean) * 100, 0)`% higher in complex food webs (table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}). This was not due to any difference between egg and larval parasitoids (fig. \ref{fig:UV_Landscape}), but was due to the lower average survival in complex food webs (contrast `r round(filter(tidy_foodweb_alphas, type=="Diff", term=="(Intercept)")$mean*-1, 2)`
[`r round(filter(tidy_foodweb_alphas, type=="Diff", term=="(Intercept)")$conf.low*-1, 2)`,
`r round(filter(tidy_foodweb_alphas, type=="Diff", term=="(Intercept)")$conf.high*-1, 2)`]).

## Food-web complexity reduces genetic constraints

\indent The curvature of the adaptive landscape indirectly affects genetic constraints and is influenced by directional, quadratic, and correlational selection gradients. Our food-web treatment did not alter correlational selection for any combination of traits (table \ref{Table:Gradients}). Similarly, our food-web treatment did not influence quadratic selection on either chamber diameter or clutch size (table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}A,C). In contrast, our food-web treatment did alter quadratic selection acting on oviposition preference (table \ref{Table:Gradients}). The negative relationship between oviposition preference and larval survival dampened at high densities in complex, but not simple food webs (fig. \ref{fig:UV_Landscape}). This dampened relationship was partly due to a trend for nonlinear selection imposed by larval parasitoids ($\gamma_{\text{Pref:Pref}}$=
`r round(filter(tidy_egglarval_grads, type=="Complex", term=="I(sc.sqrt.Pref^2)")$mean,2)` 
[`r round(filter(tidy_egglarval_grads, type=="Complex", term=="I(sc.sqrt.Pref^2)")$conf.low,2)`,
`r round(filter(tidy_egglarval_grads, type=="Complex", term=="I(sc.sqrt.Pref^2)")$conf.high,2)`]), but was also magnified by the lower average survival in complex food webs.

@@

@@

@@

Gall midges experience more selective and constraints in simple food webs.
-->

<!-- 
Natural history and comparison to other studies of natural selection in galling insects.

>>>>

\noindent 1. University of British Columbia, Department of Zoology, Vancouver, BC V6T 1Z4, Canada; <!-- consider removing -->

@@

*Keywords*: fitness landscape, host-parasitoid, natural selection, G-matrix, community context, ecological networks, evolutionary ecology, eco-evolutionary dynamics.

@@

> "What escapes the eye, however, is a much more insidious kind of extinction: the extinction of ecological interactions." @Janzen1974

@@

The adaptive landscape provides a powerful framework for understanding how natural selection has shaped the evolution of biodiversity ---from genes to phenotypes to species [@Wright1931; @Simpson1944; @Arnold2001]. More than a metaphor, the adaptive landscape links quantitative genetic and phenotypic variation to evolution by natural selection [@Lande1979; @Arnold1984applications; @Arnold1984theory]. Ecological interactions often play a key role in shaping adaptive landscapes, as evidenced by the role of antagonistic [@Schluter2000; @Abrams2000] and mutualistic [@Bronstein2006] interactions in driving evolutionary change.<!--, often play a key role in shaping the adaptive landscape and phenotypic evolution in multiple traits [; ;]. For example, competition for shared prey in threespine stickleback has driven their adaptive divergence into ecotypes with specialized body and gill morphologies [@Schluter2000, other papers]. Similarly, the elongated, red corollas of many flowering species have evolved in response to the efficient pollination of hummingbirds [cite].--> Although there is clear evidence that pairwise interactions can shape the adaptive landscape, we do not know how the adaptive landscape is shaped by the community context [@McPeek2017; @terHorst2018]<!--SOFTEN THIS STATEMENT-->. Resolution on how the community context shapes phenotypic evolution is urgently needed though, given the rapid impacts of climate change on ecological communities [@Scheffers2016].

Ecological networks, such as food webs describing who-eats-whom in ecological communities, provide an explicit representation of the community context [@Bascompte2014; @McCann2012]<!--SENTENCE SEEMS MISPLACED-->. <!--These networks describe the interdependency of populations within an ecological community, providing an effective framework for predicting how the strength and organization of ecological interactions (network structure) affect population and community dynamics [@Bascompte2014; @McCann2012]. Nevertheless, these networks often lack quantitative information on how ecological interactions influence selection on phenotypic traits. Thus, it remains unclear how network structure shapes adaptive landscapes of interacting populations. This represents a major barrier in our current ability to predict how changes in community context will affect phenotypic evolution.--> Here, we integrate ecological networks and adaptive landscapes to understand how community context constrains evolutionary change [@Arnold1992]. Different aspects of evolutionary constraints can be inferred by quantifying the slope and curvature of the adaptive landscape. For example, the slope is determined by directional selection gradients acting on each phenotypic trait and influences the trajectory of evolutionary change [@Lande1979;@Arnold1992]. Evolutionary trajectories become more constrained with an increase in the number of traits under selection, as this diminishes the number of optimal solutions [@Arnold2003]. The curvature of the adaptive landscape can also constrain evolution through its indirect effect on genetic constraints [@Arnold1992; @Hansen2008]. Genetic constraints are largely governed by a population's **G**-matrix ---the additive genetic variances and covariances between traits [@Hansen2008]. In general, genetic constraints will increase with the number of traits under directional or stabilizing selection, as this will decrease the additive genetic variance in those traits [@Hansen2008]. Genetic constraints may also increase with the number of trait combinations under correlational selection, as this type of selection decreases the evolutionary independence of traits [@Hansen2008]. If we want to predict how community context constrains evolutionary change, we must understand how ecological networks shape the adaptive landscape.
<!-- need to cite more than just Hansen and Houle 2008 here -->

Global change is simplifying the structure of ecological communities, which may influence evolutionary constraints in a number of ways. For example, in a multitrophic community or food web, if different consumers impose directional selection on different traits of a shared resource, then more complex food webs may constrain evolution by increasing the number of traits under selection. Alternatively, if consumers impose selection on different values of a trait, then their selective effects would cancel each other out in more complex food webs. To examine these different possibilities (among others), <!--To examine the effect of network structure on adaptive landscapes, -->we conducted a field experiment that manipulated the complexity of trophic interactions (hereafter food-web complexity) associated with an abundant insect herbivore (*Iteomyia salicisverruca*; Family Cecidomyiidae)(fig. \ref{fig:Conceptual}). The larvae of this herbivore induce tooth-shaped galls when they feed on the developing leaves of willow trees [*Salix* sp., @Russo2006]. These galls protect larva from attack by generalist predators (e.g. ants, spiders), but they suffer high mortality from egg and larval parasitoids [@Barbour2016]. We manipulated food-web complexity by either excluding the guild of larval parasitoids (simple food web) or allowing both egg and larval parasitoids to impose selection on gall midge traits (complex food web; fig. \ref{fig:Conceptual}). <!--Food-web complexity may influence the adaptive landscape in a number of ways. For example, if each parasitoid imposes directional selection on a different trait, then more complex food webs may constrain evolution by increasing the number of traits under selection. Alternatively, if parasitoids impose selection on different values of a trait, then their selective effects would cancel each other out in more complex food webs. To examine these different possibilities (among others), -->We then applied modern statistical methods to quantify the effect of food-web complexity on the slope and curvature of the gall midge's adaptive landscape. Taken together, our study gives insight to how local extinctions, and concomitant decreases in food-web complexity, may constrain the evolution of interacting populations.

@@

We collected data on three different traits that we expected to influence larval survival based on our previous work [@Barbour2016] and others work with gall midges [@Weis1983; @Heath2018]. First, we measured gall diameter as the size of each gall chamber to the nearest 0.01 mm at its maximum diameter (perpendicular to the direction of plant tissue growth). Our previous work has shown that larger galls are associated with higher survival [@Barbour2016]. Second, we measured clutch size by counting the number of chambers in each gall [@Weis1983; @Heath2018]. All larvae collected from the same multi-chambered gall were scored with the same clutch size. Third, we measured oviposition preference as the density of larvae observed on a plant in an independent survey. We did this by randomly sampling five branches per tree and counting the number of individual gall chambers (number of larvae). We then converted these counts to a measure of larval density per 100 shoots by counting the number of shoots on the last branch we sampled. All larvae collected from the same plant were scored with the same oviposition preference. Measuring larval densities on plants in the field is a common method for measuring oviposition preference [@Gripenberg2010]; however, caution must be taken in inferring 'preference' as larval densities can be influenced by processes other than preference [@Singer1986]. Fortunately, a couple of features of our study system suggest that larval densities may be a good proxy for oviposition preference. For example, since our data comes from a randomized placement of plant genotypes in a common garden, there is no consistent bias in which plant genotypes females are exposed to while searching for oviposition sites. Also, egg predation is a minor source of mortality for galling insects in general [@Hawkins1997]; therefore, we do not expect any prior egg predation to bias our estimates of observed larval densities.

@@

The indirect effects of selection on genetic constraints is governed by the curvature of the adaptive landscape ($\text{C}=\gamma - \beta \beta^\text{T}$), which in our study corresponds to:
<!--The curvature (**C**) of the adaptive landscape is a matrix that is determined by directional, quadratic, and correlational selection gradients, such that \boldmath$\text{C}=\gamma - \beta \beta^\text{T}$, or more explicitly:-->

@@

All analyses and visualizations were conducted in R [@R2018]. Unless otherwise noted, we report mean estimates of trait-fitness relationships and selection gradients with 95% confidence intervals in brackets. Note that for visualizing the adaptive landscape we restrict trait axes to $\pm 1$ SD of the mean trait value. This emphasizes the fact that we can only reliably estimate the shape of the adaptive landscape near the mean phenotype of the population [@Arnold2001]. We also plot mean larval survival on a natural log scale to accurately reflect the mathematical definition of the adaptive landscape [@Arnold2003].

@@

<!-- TRIPLE CHECK ALL RESULT OUTPUTS AND WRITTEN INTERPRETATIONS -->

## Simpler food webs increase selective constraints

\indent We found that more gall midge traits experienced directional selection in the simple (3 of 3) vs. complex food-web treatment (1 of 3)(table \ref{Table:Gradients}). For example, we observed directional selection for smaller clutch sizes in the simple food web, but there was no evidence of selection acting on this trait in the complex food web (table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}C). This absence of selection appeared to be a result of conflicting selection pressures imposed by each guild of parasitoids (fig. \ref{fig:UV_Landscape}). Specifically, when we subset our data to focus on differences between parasitoid guilds, we found that larval parasitoids actually impose directional selection for larger clutch sizes ($\beta_{\text{Clutch}}$=
`r round(filter(tidy_egglarval_grads, type=="Complex", term=="sc.log.Clutch")$mean,2)` 
[`r round(filter(tidy_egglarval_grads, type=="Complex", term=="sc.log.Clutch")$conf.low,2)`,
`r round(filter(tidy_egglarval_grads, type=="Complex", term=="sc.log.Clutch")$conf.high,2)`]). In the simple food web, we also observed clear evidence of directional selection for midges to avoid ovipositing on plants with high densities of conspecifics (table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}B); however, this relationship was weaker in complex food webs (table \ref{Table:Gradients}). This was likely a result of larval parasitoids imposing greater mortality on egg parasitoids at high gall midge densities (see Selection on the extended phenotype of egg parasitoids section), and thus a less than additive effect on gall midges. Chamber diameter experienced positive directional selection in both food-web treatments, but selection was `r round(((filter(tidy_foodweb_grads, type=="Complex", term=="sc.Diam")$mean - filter(tidy_foodweb_grads, type=="Simple", term=="sc.Diam")$mean) / filter(tidy_foodweb_grads, type=="Simple", term=="sc.Diam")$mean) * 100, 0)`% higher in complex food webs (table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}). This was not due to any difference between egg and larval parasitoids (fig. \ref{fig:UV_Landscape}), but was due to the lower average survival in complex food webs (contrast `r round(filter(tidy_foodweb_alphas, type=="Diff", term=="(Intercept)")$mean*-1, 2)`
[`r round(filter(tidy_foodweb_alphas, type=="Diff", term=="(Intercept)")$conf.low*-1, 2)`,
`r round(filter(tidy_foodweb_alphas, type=="Diff", term=="(Intercept)")$conf.high*-1, 2)`]).

## Simpler food webs increase genetic constraints

\indent The curvature of the adaptive landscape indirectly affects genetic constraints and is influenced by directional, quadratic, and correlational selection gradients. Our food-web treatment did not alter correlational selection for any combination of traits (table \ref{Table:Gradients}). Similarly, our food-web treatment did not influence quadratic selection on either chamber diameter or clutch size (table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}A,C). In contrast, our food-web treatment did alter quadratic selection acting on oviposition preference (table \ref{Table:Gradients}). The negative relationship between oviposition preference and larval survival dampened at high densities in complex, but not simple food webs (fig. \ref{fig:UV_Landscape}B). This dampened relationship was partly due to a trend for nonlinear selection imposed by larval parasitoids ($\gamma_{\text{Pref:Pref}}$=
`r round(filter(tidy_egglarval_grads, type=="Complex", term=="I(sc.sqrt.Pref^2)")$mean,2)` 
[`r round(filter(tidy_egglarval_grads, type=="Complex", term=="I(sc.sqrt.Pref^2)")$conf.low,2)`,
`r round(filter(tidy_egglarval_grads, type=="Complex", term=="I(sc.sqrt.Pref^2)")$conf.high,2)`]), but was also magnified by the lower average survival in complex food webs.

@@

<!-- CONSIDER AVOIDING EXTENDED PHENOTYPE AND FOCUS ON INDIRECT EFFECTS OR INDIRECT SELECTION -->

@@

<!--Summary-->
Our key finding was that phenotypic evolution in gall midges was more constrained in simpler food webs. More traits contributed to the slope of the adaptive landscape in the simpler food web, suggesting that the trajectory of phenotypic evolution is more constrained. Similarly, the simple food-web treatment indirectly increased the number of genetic constraints, which could act to constrain the gall midge's adaptive potential in the face of novel selection pressures. We also found evidence of indirect selection pressures in more complex food webs, suggesting that the adaptive landscape may be more dynamic in complex food webs. Taken together, our study provides experimental evidence from the field that the simplification of ecological communities may constrain the adaptive potential of remaining populations.

<!--Selective constraints-->
All three phenotypic traits we examined experienced directional selection in the simpler food web, indicating that there is an optimal phenotype that maximizes larval survival (i.e. large chamber diameter, small clutch size, and avoidance of conspecifics). In contrast, we did not observe selective constraints on clutch size and oviposition preference in the more complex food web. This suggests that there are many optimal phenotypes (adaptive peaks), or in this case, a flat adaptive landscape where there are no fitness consequences for phenotypic change in these traits. This also implies that as selective constraints dampen in more complex food webs, then the trajectory of evolution becomes more determined by genetic constraints. This is made clear if we inspect the 'Lande equation', $\Delta z = \text{G} \beta$, where G corresponds to the population's G-matrix. As the components of $\beta$ go to zero, then $\Delta z$ is more influenced by the G-matrix. Interestingly, this corresponds to @Schluter1996's hypothesis that phenotypic evolution often follows 'genetic lines of least resistance'. @Schluter1996 found support for this hypothesis from data on natural populations of several vertebrate species, including threespine sticklebacks, a few species of songbirds, and mice from the genus *Peromyscus*. All of these species occupy intermediate trophic levels and are likely embedded in complex food webs, which is consistent with our suggestion that genetic constraints may have a stronger influence in more complex food webs.

<!--Indirect effects of selective constraints-->
We also found evidence for more genetic constraints in simpler food webs due to indirect effects of selection on the population's G-matrix. The ability of a population to adapt to novel selection pressures (evolvability) is largely governed by the structure of its G-matrix [@Hansen2008]. When selection favors genetic covariance between traits (positive or negative), this results in less autonomy of evolutionary responses to changing environments. Similarly, decreases in genetic variance constrain potential for the trait itself to evolve. Together, this suggests that simpler food webs may decrease the evolutionary potential of populations by indirectly selecting for decreases in genetic variance in multiple traits and favoring trait integration. Current theory often assumes genetic variances and covariances remain constant over time rather than dynamically changing with the community context [@McPeek2017; @Guimaraes2017]. Our empirical results highlight the need to explore the evolutionary consequences of not only direct effects of selection, but indirect effects on genetic constraints that are shaped by the community context.

<!-- Generality of results -->
The generality of our results likely depends on the relative abundance and functional differences between consumers in a community. In the simplest case, when consumers do not differ from each other, then we do not expect changes in food-web complexity to modify selective constraints. Also, many consumers may be at too low of abundances to impose selection on their resources. Rank abundance curves [@Preston1948] and the disproportionate number of weak interactions in diverse communities [@Paine1992] support this notion. When consumers are abundant though, the effect of food-web complexity will depend on whether different species impose conflicting selection pressures or select for distinct traits. For example, parisitoids and birds impose conflicting selection pressures on the size of galls induced by the fly *Eurosta solidaginis* [@Weis1985; @Abrahamson1997; @Start2016]. Recent studies in this system have shown that decreases in the relative abundance of birds, due to either small patch sizes [@Start2016] or proximity to urban areas [@Start2018] causes a shift from neutral to directional selection on gall size. On the other hand, different consumers may impose selection on different traits, favoring trait integration and increasing genetic constraints. Examples of this include strong genetic covariances in plant resistance to different insect herbivores [@Maddox1990; @Wise2007; @Wise2013], although there are also examples where these covariances are weak [@Roche1997; @Barbour2015], or vary from year-to-year [@Johnson2007]. We suggest that gaining predictive insight to the effects of food-web disassembly requires an understanding of the mechanisms governing the assembly of trophic interactions.

<!--Indirect evolutionary effects-->
Our results suggest that the simplification of ecological communities may not only directly affect connected species, but also result in indirect evolutionary effects. In our study, this indirect effect arises from egg parasitoids being released from intraguild predation in simpler food webs. This release occurs more on trees with high larval densities, which could intensify selection on gall midge oviposition preference. A growing number of experiments over the past two decades have demonstrated the presence and potential importance of indirect evolutionary effects that emerge in a community context [@Juenger1998; @Stinchcombe2001; @Lankau2007; @Walsh2008; @Walsh2010; @terHorst2010; @Sahli2011; @Lau2012; @terHorst2015; @Schiestl2018]<!--; Start et al. in press-->. If indirect evolutionary effects are common [@Walsh2013], then predicting evolutionary trajectories resulting from the simplification of food webs will require evolutionary studies to explicitly account for the ecological networks that species are embedded in.

<!-- Community context shaping adaptation to environmental change --> 
There are a growing number of theoretical studies on adaptation to directional environmental change that incorporate species interactions [e.g. @deMazancourt2008; @Johansson2008; @Norberg2012; @Osmond2017PredatorsHelpPrey]. This work has given insight to the mechanisms by which pairwise interactions modify the adaptive potential of species. Our study hints at a novel mechanism, whereby food-web complexity flattens the adaptive landscape, thus facilitating future adaptation by allowing genetic and phenotypic variation to persist. This mechanism only emerges once we move beyond pairwise interactions to consider selection on multiple traits in a community context. However, a study on competition has highlighted that we may expect the opposite effect of species diversity in competitive communities [@deMazancourt2008]. This negative effect of species diversity on adaptation occurs because there is a greater chance that the optimal phenotype is already occupied by a competitor in a more diverse community. More work is needed to examine how the distribution of different interaction types affects adaptation to environmental change in species-rich communities.

<!-- Conclusions -->
Our study gives insight to how local extinctions, and concomitant decreases in food-web complexity, increase evolutionary constraints. This hints at a potential insidious effect of local extinctions that compromises the robustness of remaining populations to future environmental change. Our work also highlights some key challenges for predicting phenotypic evolution in the face of global change. First, the simplification of ecological communities may actually reduce the predictability of phenotypic evolution, because knowledge is required of both selective and genetic constraints, rather than potentially just genetic constraints in more complex systems. Second, many theoretical models of eco-evolutionary dynamics focus on phenotypic change in a single trait, yet our results highlight that the number of traits under selection may change with the community context. Importantly, while the net effects of selection in complex communities may be dampened, we still found that different species/guilds imposed different selection pressures. Knowing these hidden selection pressures is critical for prediction, because the trajectory of evolution will depend on the nature of change in the community context. We expect that a continued integration of adaptive landscapes and ecological networks will enhance our ability to predict the evolutionary consequences of changes in ecological communities.

@@

Gall midges experience more selective and constraints in simple food webs.
-->
<!--More phenotypic traits experienced directional selection in the simple food-web treatment, suggesting that the trajectory of phenotypic evolution is more constrained.-->

<!--, because the egg parasitoids that attack larva on high density trees have a higher chance of surviving.-->

<!--In the context of global change, our results imply that the simplification of ecological communities may actually reduce the predictability of phenotypic evolution of remaining populations. This is because knowledge is required of both selective and genetic constraints, rather than potentially just genetic constraints in more complex systems.-->

<!-- what should the focus be? 
Conceptually, I don't think it'll be that interesting to try and review past studies.

I think it would be more interesting to discuss the implications of our results and there generality. In other words, under what conditions do we expect them to hold. When will they break down?

Highlight that we could gain more predictive power from theoretical models that also examine how selection modifies selection pressures of consumers. i.e. we need models of ecological character displacement that go beyoned pairwise interactions. In some communities, intraguild predation will be common so we also need models that examine how the community context shapes the evolution of trophic levels and the resulting feedbacks on associated communities (cite JP's Frontiers in Ecology and Evolution MS).

Such indirect effects may be quite common when moving from complex to simple communities, and highlights the fact that species evolve in a community context.<!-- and the release of selection pressures may have rippling effects.-->  <!--If indirect evolutionary effects are common [@Walsh2013], then predicting the evolutionary trajectories of remaining populations after local extinctions will not only require knowledge of these indirect effects, but also require knowledge on the order of local extinctions.

It is well known that increases in consumer diversity can driver character divergence in foraging traits due to ecological character displacment [@Schluter2000]. In this case, we would expect the local extinction of a consumer to 
- Ecology example: selection on gall size in Eurosta
- Evolution example: ecological character displacement tends to drive divergence.
-->

<!--
Detailed case studies of natural selection on a single trait in a community context highlight the potential generality of our results. For example, galls induced by the fly *Eurosta solidaginis* experience directional selection for larger sizes by parasitoids (as we observed in our study), but selection for smaller sizes by birds [@Weis1985; @Abrahamson1997; @Start2016]. Therefore, the dynamics of natural selection on gall size depend on the relative abundance of these consumers. A recent study in this system demonstrated that the loss of birds across a gradient of urbanization caused a shift from neutral to directional selection on gall size [@Start2018]. This observational study supports our experimental demonstration that the loss of consumer diversity increases selective constraints.

In a different system, the direction of selection on swimming speed in *Enallagama* damselflies depends on whether they occur in fish or fishless lakes [@McPeek1996; @McPeek1997]. Lakes with fish select for damselflies , whereas dragonflies become the top predator in fishless lakes and select for adaptations to increase swimming speed.

fish and dragonflies impose opposite directional selection on swimming speed in *Enallagma* damselflies . This has resulted in

Lakes with fish likely have a much more diverse community of damselfly predators (multiple fish species and some smaller dragonfly species),

Our study focused on an experimental manipulation of a focused community of consumers, leaving an open question to how our results generalize to more diverse consumer communities.

-->
<!--
The generality of our results depends on the extent to which different consumer species are functionally different and their relative abundance. In the simplest case, when consumers do not differ from each other (i.e. neutral communities), then we do not expect changes in food-web complexity to modify selective constraints. Also, many consumer species may be at too low of abundances --><!--This may be the case with many consumers in diverse communities, where their abundances are simply not high enough --><!-- to impose selection on the mean trait values of resource populations. Rank abundance curves [@Preston1948] and the disproportionate number of weak interactions in diverse communities [@Paine1992] support this notion. For abundant species though, the effect of food-web complexity depends on whether different species impose selection on different trait values vs. distinct traits.

Current theory on character displacement suggests that consumers will usually diverge in foraging traits when foraging on a biological resource. On one end, theory should be extended beyond two consumers to examine trait dynamics in more diverse consumer communities. On the other hand, we also need to examine selection in the resources themselves. 
-->
<!--
Actually, different predator species impose selection on different axes of trait variation. 
For example, habitat shifts into lakes where dragonflies are the top predator (instead of fish) promote adaptations for increased swimming speed in multiple lineages of *Enallagma* damselflies [@McPeek1996].
Phenotypic selection study with damselflies [@McPeek1997].
Biochemical evolution study with damselflies [@McPeek1999], still related to swimming speed.
In McPeek's study though, there aren't dragonflies in lakes with fish?

Lakes with fish are the ancestral state, and they move to fishless lakes. Although the fishless lakes may have simpler communities, we see a clear effect of the dominance hierarchy which highlights the importance of abundance. In this case, we have a strong intraguild predator that reduces the density of intraguild prey (dragonfly larvae), thus reducing their selection pressures. This indicates that its' not simply the presence/absence of different predators, but the evenness of these selection pressures.
-->
<!--
Maybe they break down with plant-herbivore interactions? Or maybe these provide a useful example for thinking about when they'll break down or for highlighting indirect effects on genetic constraints?

We suggest that being able to predict how changes in community context shape evolutionary constraints requires a deep understanding of the ecological and evolutionary feedbacks that shape trait variation and community structure.

Here, we focus on how changes in food-web complexity affects evolutionary constraints on a single resource population. However, predicting h
